---
title: "A1. salmonMSE simple model"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE
)

library(salmonMSE)
library(here)
```

# Introduction to operating model

The exercise below is an introduction to the basic mechanics of working with salmonMSE.

The operating model is an object where we put parameters describing the life history of our case study and the management strategy to be evaluated.

In R, the operating model is an S4 object of class `SOM`. We can view the documentation with:

```{r eval=FALSE}
library(salmonMSE)
class?SOM
```

The `SOM` object is built out of constituent objects that organizes the inputs and different management levers. Creating a full `SOM` looks something like this:

```{r eval=FALSE}
# Note: demonstration only, this code chunk will not work
SOM <- new(
  "SOM",
  Bio = Bio,
  Habitat,
  Hatchery,
  Harvest,
  Historical
)
```

We'll work with 5 smaller objects, `Bio`, `Habitat`, `Hatchery`, `Harvest`, and `Historical`, each of which has their own documentation page. For example:

```{r eval=FALSE}
class?Bio
```

While we need to create all five objects, we may carefully work with as few as 1-2 objects depending on the case study. But let's start simple.

# Simple case study

## Bio object

The `Bio` object component of the operating model describes the basic biology (maturity, lifespan) of the population. 
In general, we need to describe what happens in each age class.

Here's the first example:

```{r}
nsim <- 20
SAR <- 0.01

set.seed(4389)
kappa_mean <- 3
kappa_sd <- 0.5

library(EnvStats)
kappa <- rnormTrunc(nsim, kappa_mean, kappa_sd, min = 1, max = Inf)

Bio <- new("Bio")
Bio@Name <- "Prototype"
Bio@maxage <- 3
Bio@p_mature <- c(0, 0, 1)
Bio@SRrel <- "Ricker"
Bio@kappa <- kappa
Bio@Smax <- 1000
Bio@Mjuv_NOS <- c(0, -log(SAR))
Bio@fec <- c(0, 0, 5000)
Bio@p_female <- 0.5
```

Alternatively, the `Bio` object can be filled in one call:

```{r eval=FALSE}
Bio <- new(
  Class = "Bio",
  maxage = 3,
  p_mature = c(0, 0, 1),
  SRrel = "Ricker",
  kappa = kappa,
  Smax = 1000,
  Mjuv_NOS = c(0, -log(SAR)),
  fec = c(0, 0, 5000),
  p_female = 0.5
)
```

The information contained in this object:

- The lifespan is 3 years
- All fish mature at age 3
- Natural-origin juvenile marine survival is 0.01, occurring between age 2 to 3
- 50% of spawners are female, each with fecundity of 5 thousand eggs
- The density-dependence of the population is described with a Ricker stock-recruit function. The productivity parameter (kappa) is stochastic over 20 simulations with an average of 3 adults/spawner. The Smax parameter is 1,000 spawners.

There's also some other default parameters that we are not discussing (for example, en-route survival to spawning grounds is 1, see `class?Bio`).

## Harvest object

In the `Bio` object, we have egg-juvenile survival described in the Ricker stock recruit relationship, and juvenile marine survival is 0.01.

Next, we describe the exploitation rate in the fishery in the `Harvest` object:

```{r}
Harvest <- new(
  "Harvest",
  u_preterminal = 0,
  u_terminal = 0.2,
  vulT = c(1, 1, 1)
)
```

The information in this object:

- There is no preterminal marine fishery exploitation 
- The terminal exploitation rate is 0.2. This corresponds to the maximum rate experienced in any age class
- The terminal fishery selects all ages equally (vulnerability is identical for all ages)

## Other objects

We have three more objects to fill:

- We don't have a hatchery
- We don't model freshwater life stages
- We will simulate with 1,000 individuals at the start of the projection

```{r}
# Set releases to zero to turn off hatchery
Hatchery <- new(
  "Hatchery",
  n_yearling = 0,
  n_subyearling = 0
)

# Use flag to be explicit about assumptions
Habitat <- new(
  "Habitat",
  use_habitat = FALSE
)

# Set number of juveniles at beginning of simulation
Historical <- new(
  "Historical",
  InitNjuv_NOS = 1000,
  InitNjuv_HOS = 0
)
```

## Run projection

We create the `SOM` object now that we have individual objects completed:

```{r}
proyears <- 50
SOM <- new(
  "SOM",
  nsim = nsim,
  proyears = proyears,
  Bio = Bio,
  Habitat = Habitat,
  Hatchery = Hatchery,
  Harvest = Harvest,
  Historical = Historical
)
```

We run the projection with a call to `salmonMSE()`. 

Generally, it's a good idea to save the output, especially for larger models. 

```{r run-sim, eval=TRUE}
SMSE <- salmonMSE(SOM)

if (!dir.exists(here("output"))) dir.create(here("output"))
saveRDS(SMSE, here(file.path("output", "A1-simple.rds")))
```

# Reporting

We then generate a markdown report of the projection with `report()`:

```{r, echo=FALSE}
SMSE <- readRDS(here(file.path("output", "A1-simple.rds")))
```

```{r report, eval=TRUE}
report(
  SMSE, 
  name = "Simple case study", 
  filename = "A1-simple", 
  dir = here("output")
)
```

## Exploring output

The output (class `SMSE`) from the projection contains various arrays of the state variables during the projection:

```{r eval=FALSE}
class?SMSE
SMSE@NOS
```

```{r}
slotNames(SMSE)
```

For example, in slot `NOS`, we have an array by simulation x stock x age x year that gives us the natural-origin spawners:

```{r}
dim(SMSE@NOS)
```

## Figures

There are plotting functions to help us explore the output. 

`plot_statevar_ts()` generates a time series figure with base graphics. By default, it generates line plots for each stochastic simulation in the projection:

```{r}
plot_statevar_ts(SMSE, "NOS")
```

Adding an additional argument gives us a summary quantile by year:

```{r}
plot_statevar_ts(SMSE, "NOS", quant = TRUE)
```

We can also explore the distribution of outcomes within an individual year. `plot_statevar_hist()` returns a histogram. By default, it plots the values in the last year of the projection:

```{r}
plot_statevar_hist(SMSE, "NOS")
```

And again, an additional argument gives us control to plot distributions in other projection years:

```{r}
plot_statevar_hist(SMSE, "NOS", y = 1)
```

All these figures return results from a single model run. We will explore how to compare model runs later on.
