---
title: "C1. Running multiple scenarios"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE
)
library(salmonMSE)
library(here)
```

# Introduction

Here, we demonstrate how to organize a simulation that evaluates alternative management strategies across uncertain biological parameters. 

The management strategy to be evaluated is the terminal exploitation rate (ER), which is evaluated in separate simulation runs.

The uncertain parameters for which we want to evaluate robustness is marine survival (SAR) as well as the stock-recruit (SR) Ricker parameters.
We incorporate uncertainty in Ricker parameters within operating model, while marine survival is specified in separate operating models.

Each operating model is a unique description of the system dynamics. Thus, we will run a separate operating model for each combination of ER and marine survival.

I find the best approach is to write a single model run, then use that code in a wrapper function. 
The wrapper function can then used to repeat the simulation for combination of parameters.
Writing a wrapper function also facilitates the vectorization capabilities of R using the `lapply()` function to repeat over parameter combinations as well as the parallel processing capabilities.

# Single model run

Below is a single code chunk for a single model projection:

```{r single-run}
library(salmonMSE)
nsim <- 20
set.seed(234)

SAR <- 0.01
ER <- 0.1

# We are sampling Ricker alpha values which represent maximum egg-smolt survival
# While salmonMSE expects productivity values (recruits/spawner) in the kappa slot, 
# we can set the phi conversion term to 1 to directly enter alpha
alpha_mean <- 0.1
alpha_sd <- 0.15
alpha <- rlnorm(nsim, log(alpha_mean) - 0.5 * alpha_sd^2, alpha_sd)

# We are sampling Ricker Emax = 1/beta (egg production that maximizes the stock-recruit function)
# represent maximum egg-smolt survival
# While salmonMSE expects Smax, 
# we can set both phi and tau conversions term to 1 to directly enter Emax
# Here mean Emax is the egg production from 1000 spawners (fecundity = 5,000 eggs)
Emax_mean <- 1000 * 5000
Emax_sd <- 0.2
Emax <- rlnorm(nsim, log(Emax_mean) - 0.5 * Emax_sd^2, Emax_sd)

Bio <- new(
  Class = "Bio",
  maxage = 3,
  p_mature = c(0, 0, 1),
  SRrel = "Ricker",
  kappa = alpha,
  Smax = Emax,
  phi = 1,
  tau = 1,
  Mjuv_NOS = c(0, -log(SAR)),
  fec = c(0, 0, 5000),
  p_female = 0.5
)

Harvest <- new(
  "Harvest",
  u_preterminal = 0,
  u_terminal = ER,
  vulT = c(1, 1, 1)
)

Hatchery <- new(
  "Hatchery",
  n_yearling = 0,
  n_subyearling = 0
)

# Use flag to be explicit about assumptions
Habitat <- new(
  "Habitat",
  use_habitat = FALSE
)

# Set number of juveniles at beginning of simulation
InitNjuv_NOS <- array(1000, c(nsim, Bio@maxage, 1))
InitNjuv_NOS[, 1:2, ] <- 1000 / SAR
Historical <- new(
  "Historical",
  InitNjuv_NOS = InitNjuv_NOS,
  InitNjuv_HOS = 0
)

proyears <- 40
SOM <- new(
  "SOM",
  nsim = nsim,
  proyears = proyears,
  Bio = Bio,
  Habitat = Habitat,
  Hatchery = Hatchery,
  Harvest = Harvest,
  Historical = Historical
)
```


```{r run-sim1, eval=FALSE}
SMSE <- salmonMSE(SOM)

if (!dir.exists(here("output"))) dir.create(here("output"))
saveRDS(SMSE, here(file.path("output", "C1-one-run.rds")))
```


# Wrapper function

Now let's put the above code chunk into a function. We'll call it `wrapper()`.

This function has three arguments:

- `SAR_vector` that contains various marine survival values
- `ER_vector` that contains various exploitation rate values, same length as `SAR_vector`
- `i` which is an integer that indexes both `SAR_vector` and `ER_vector`. When `wrapper()` is called, you supply a value for `i` and the function will grab the `i`-th value in `SAR_vector` and `ER_vector` for the model run.

The rest of the function should be the same as the previous chunk.

It will re-create the same operating model each time the function is called. Each model is identical except for the SAR and ER value.

The wrapper function will then also run the projection:

```{r}
wrapper <- function(i, SAR_vector, ER_vector) {

  require(salmonMSE)
  
  SAR <- SAR_vector[i]
  ER <- ER_vector[i]

  nsim <- 20
  set.seed(234)

  # We are sampling Ricker alpha values which represent maximum egg-smolt survival
  # While salmonMSE expects productivity values (recruits/spawner) in the kappa slot, 
  # we can set the phi conversion term to 1 to directly enter alpha
  alpha_mean <- 0.1
  alpha_sd <- 0.15
  alpha <- rlnorm(nsim, log(alpha_mean) - 0.5 * alpha_sd^2, alpha_sd)
  
  # We are sampling Ricker Emax = 1/beta (egg production that maximizes the stock-recruit function)
  # represent maximum egg-smolt survival
  # While salmonMSE expects Smax, 
  # we can set both phi and tau conversions term to 1 to directly enter Emax
  # Here mean Emax is the egg production from 1000 spawners (fecundity = 5,000 eggs)
  Emax_mean <- 1000 * 5000
  Emax_sd <- 0.2
  Emax <- rlnorm(nsim, log(Emax_mean) - 0.5 * Emax_sd^2, Emax_sd)

  Bio <- new(
    Class = "Bio",
    maxage = 3,
    p_mature = c(0, 0, 1),
    SRrel = "Ricker",
    kappa = alpha,
    Smax = Emax,
    phi = 1,
    tau = 1,
    Mjuv_NOS = c(0, -log(SAR)),
    fec = c(0, 0, 5000),
    p_female = 0.5
  )

  Harvest <- new(
    "Harvest",
    u_preterminal = 0,
    u_terminal = ER,
    vulT = c(1, 1, 1)
  )

  Hatchery <- new(
    "Hatchery",
    n_yearling = 0,
    n_subyearling = 0
  )

  # Use flag to be explicit about assumptions
  Habitat <- new(
    "Habitat",
    use_habitat = FALSE
  )

  # Set number of juveniles at beginning of simulation
  InitNjuv_NOS <- array(1000, c(nsim, Bio@maxage, 1))
  InitNjuv_NOS[, 1:2, ] <- 1000 / SAR
  Historical <- new(
    "Historical",
    InitNjuv_NOS = InitNjuv_NOS,
    InitNjuv_HOS = 0
  )

  proyears <- 20
  SOM <- new(
    "SOM",
    nsim = nsim,
    proyears = proyears,
    Bio = Bio,
    Habitat = Habitat,
    Hatchery = Hatchery,
    Harvest = Harvest,
    Historical = Historical
  )
  
  SMSE <- salmonMSE(SOM)

  return(SMSE)
}
```

# Creating simulation design

One more thing, we also want to create our simulation design, i.e., which ERs and marine survival values we will use.

Here's one way to create a grid of values with `expand.grid()`. Each row will identify a combination of values for our simulation:

```{r}
simulation_design <- expand.grid(
  SAR_vector = seq(0.005, 0.03, 0.005),
  ER_vector = seq(0.2, 0.7, 0.1)
)
```

# Running wrapper function

We have our wrapper function and simulation design, we can now run the simulation.

There's various ways to run the projections depending on our computation capabilities, e.g., old laptop, new desktop, etc.

For example, we can run a loop over each row of our `simulation_design` data frame or use `lapply()` to facilitate repeated simulations. 

```{r, eval=FALSE}
# Not intended to run, run in parallel instead?
SMSE_list <- list()
for (i in 1:nrow(simulation_design)) {
  SMSE_list[[i]] <- wrapper(
    i, 
    SAR_vector = simulation_design$SAR_vector,
    ER_vector = simulation_design$ER_vector
  )
}
```

```{r, eval=FALSE}
# Not intended to run, run in parallel instead?
SMSE_list <- lapply(
  1:nrow(simulation_design), 
  wrapper, 
  SAR_vector = simulation_design$SAR_vector,
  ER_vector = simulation_design$ER_vector
)
```

But loops are computationally slower compared to parallel processing. We use the `snowfall` package to run simulations over 10 CPUs, then save the results:

```{r run-sim-list, eval=TRUE}
library(snowfall)
cpus <- parallel::detectCores() - 2
sfInit(parallel = TRUE, cpus = cpus)

SMSE_list <- sfLapply(
  1:nrow(simulation_design), 
  wrapper,
  SAR_vector = simulation_design$SAR_vector, 
  ER_vector = simulation_design$ER_vector
)

sfStop()

if (!dir.exists(here("output"))) dir.create(here("output"))
saveRDS(SMSE_list, here(file.path("output", "C1-list.rds")))
```

# Results

Now we have a list of `SMSE` objects in `SMSE_list`. 

## Compare time series

Whereas we saw previously that `report()` generates a Markdown report from a single model run, we alternatively now want to compare several model runs simultaneously in a figure.

We have the `compare()` function which generates another markdown report that plots state variables from separate model runs. 

We have 36 scenarios, so we may not be able to neatly all into individual figures. 
So let's compare just the first 3 model runs:

```{r, echo=FALSE}
SMSE_list <- readRDS(here(file.path("output", "C1-list.rds")))
```

```{r}
# Create names for each model run by pasting the SAR and ER values
SMSE_names <- paste0("SAR = ", simulation_design$SAR_vector, ", ER = ", simulation_design$ER_vector)
```

```{r, eval=TRUE}
compare(
  SMSE_list[1:3],
  SMSE_names[1:3],
  filename = "C1_compare",
  dir = here("output")
)
```


## Plotting functions

`compare_statevar_ts()` generates a time series figure with base graphics. By default, it generates line plots for each stochastic simulation in the projection (which is messy), so let's directly proceed to quantile figures:

```{r}
compare_statevar_ts(SMSE_list[1:3], var = "NOS", quant = TRUE, names = SMSE_names[1:3])
```

We can also look to distributions within scenarios with `compare_statevar_hist()`, where there are choices to plot using the `density()` or `hist()` functions:

```{r}
compare_statevar_hist(SMSE_list[1:3], var = "NOS", names = SMSE_names[1:3], type = "density")
```

```{r}
compare_statevar_hist(SMSE_list[1:3], var = "NOS", names = SMSE_names[1:3], type = "hist")
```

## Performance measures and decision tables

Risk evaluation typically entails calculating a summary metric that describes the frequency that an objective was met in the simulation. 

For example, we can calculate the probability that spawner abundance exceeded 0.85 SMSY (spawners at MSY).
We have a function `SMSY85()` that does the calculation for us:

Using the vectorization capabilities in R, we use `sapply()` to call `SMSY85()` for each of the objects in `SMSE_list`. We add an extra argument `Yrs = c(20, 20)` that specify the year range (we do the calculation at the end of the projection in year 20:

```{r}
sim_results <- simulation_design
sim_results$spawner_objective <- sapply(SMSE_list, SMSY85, Yrs = c(20, 20))
```

We want to explore the decision space: at what exploitation rates are we likely to reach our objective across plausible values of marine survival?

`plot_decision_table()` helps us plot a grid of values using ggplot.
The colour scheme maps 0-1 with red to green (0.5 is white).

We can see a nice frontier line that delineates our decision-making space: we can meet our spawner objective with higher exploitation rates only if marine survival increases as well.

Exploitation rates of 0.2 would be most robust to the marine survival scenarios that we simulated.

We can envision running a larger grid to map out the full decision space. 

```{r}
plot_decision_table(
  x = sim_results$SAR_vector,
  y = sim_results$ER_vector,
  z = sim_results$spawner_objective,
  xlab = "Marine survival",
  ylab = "Exploitation rate",
  title = "Probability of exceeding spawner objective"
)
```

Alternatively, `plot_decision_table2()` gives us more flexibility in the colour scheme by binning the summary metric into separate colour bins. Using the above figure, we refine the decision table based on the values we see in our results.


```{r}
plot_decision_table2(
  x = sim_results$SAR_vector,
  y = sim_results$ER_vector,
  z = sim_results$spawner_objective,
  xlab = "Marine survival",
  ylab = "Exploitation rate",
  zlab = "Probability\nspawner objective",
  bin = c(0, 0.6, 0.7, 0.8),
  bin_labels = c("0.00 - 0.59", "0.60 - 0.69", "0.70 - 0.79", "0.80 - 1"),
  bin_col = c("pink", "white", "green", "green4")
)
```

## Tradeoffs

Frequently, objectives are conflicting. For example, lower exploitation rates could allow for more escapement at the cost of catch. Tradeoff figures allow us to compare two summary metrics simultaneously.

First, let's calculate our second metric, average catch, with a custom function:

```{r}
Catch_fn <- function(SMSE) mean(SMSE@KT_NOS[, 1, 20])
sim_results$Mean_catch <- sapply(SMSE_list, Catch_fn)
```

Next, we can use `plot_tradeoff()` to plot the two metrics (average catch and spawner probability) against each other.

Typically, we want a nice frontier line that allows us to see the impact of our management options as in the decision tables above, but the tradeoff figure below is quite messy. 

```{r}
plot_tradeoff(
  pm1 = sim_results$spawner_objective,
  pm2 = sim_results$Mean_catch,
  x1 = sim_results$ER_vector,
  x2 = sim_results$SAR_vector,
  xlab = "Probability spawner objective",
  ylab = "Mean catch",
  x1lab = "Exploitation rate",
  x2lab = "Marine survival"
)
```

Perhaps we can subset by marine survival can improve it a bit. 

For a given value of marine survival, we see the long-term spawner-catch performance a bit more clearly with respect to exploitation rate.

```{r}
sim_subset <- sim_results[sim_results$SAR_vector == 0.02, ]

library(ggplot2)
plot_tradeoff(
  pm1 = sim_subset$spawner_objective,
  pm2 = sim_subset$Mean_catch,
  x1 = sim_subset$ER_vector,
  xlab = "Probability spawner objective",
  ylab = "Mean catch",
  x1lab = "Exploitation rate"
) +
  scale_colour_viridis_c() +
  labs(title = "Marine survival = 0.02")
```

Another alternative is to plot the decision tables next to each other:

```{r}
g1 <- plot_decision_table(
  x = sim_results$SAR_vector,
  y = sim_results$ER_vector,
  z = sim_results$spawner_objective,
  xlab = "Marine survival",
  ylab = "Exploitation rate",
  title = "Probability spawner objective"
)

g2 <- plot_decision_table(
  x = sim_results$SAR_vector,
  y = sim_results$ER_vector,
  z = round(sim_results$Mean_catch),
  xlab = "Marine survival",
  ylab = "Exploitation rate",
  title = "Mean catch"
)

cowplot::plot_grid(g1, g2)
```

