---
title: "C2. Running multiple scenarios (hatchery programs and more)"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE
)
library(salmonMSE)
library(here)
```

# Introduction

Here, we demonstrate how to organize a simulation that evaluates alternative management strategies across uncertain biological parameters. 

The management levers we want to evaluate are:

1. Hatchery production level (target releases)
2. Egg incubation survival (if we can improve survival by reducing fine sedimentation)
3. Mark selective fishing (on or off) in the terminal fishery

The uncertain parameter for which we want to evaluate robustness is age-1 marine survival, incorporated stochastically within each operating model.
Egg-fry survival is also stochastic although the mean is constant.

Each operating model is a unique description of the system dynamics. Thus, we will run a separate operating model for each combination of our management levers.

I find the best approach is to write a single model run, then use that code in a wrapper function. 
The wrapper function can then used to repeat the simulation for combination of parameters.
Writing a wrapper function also facilitates the vectorization capabilities of R using the `lapply()` function to repeat over parameter combinations as well as the parallel processing capabilities.

# Single model run

Below is a code chunk to run a single model:

```{r single-run}
library(salmonMSE)

nsim <- 20
proyears <- 30
maxage <- 3

set.seed(234)

surv1 <- 0.01
surv2 <- 0.8

Bio <- new(
  Class = "Bio",
  maxage = maxage,
  p_mature = c(0, 0.3, 1),
  #SRrel = "Ricker",
  #kappa = c(3, 4, 6),
  #Smax = 1000,
  #phi = 1,
  #tau = 1,
  Mjuv_NOS = c(-log(surv1), -log(surv2)),
  fec = c(0, 2000, 5000),
  p_female = 0.5,
  s_enroute = 0.95
)

Harvest <- new(
  "Harvest",
  u_preterminal = 0.4,
  u_terminal = 0.2,
  vulPT = c(0, 0.4, 1),
  vulT = c(0, 0.4, 1),
  MSF_PT = FALSE,
  MSF_T = FALSE
)

Hatchery <- new(
  "Hatchery",
  n_r = 1,
  n_subyearling = 100000,
  s_prespawn = 0.95,
  s_egg_subyearling = 0.7,
  Mjuv_HOS = Bio@Mjuv_NOS,
  p_mature_HOS = Bio@p_mature,
  gamma = 0.8,
  m = 1,
  pmax_esc = 1,
  pmax_NOB = 0.8,
  ptarget_NOB = 0.75,
  phatchery = NA,
  fec_brood = Bio@fec,
  fitness_type = c("none", "none")
)

sdev_sd <- 0.1

set.seed(84)
fry_sdev <- EnvStats::rnormTrunc(nsim * proyears, 1, sdev_sd, min = 0, max = 1/0.7) |> 
  matrix(nsim, proyears)

Habitat <- new(
  "Habitat",
  use_habitat = TRUE,
  egg_rel = "BH",
  egg_prod = 0.21,
  egg_capacity = 1000 * 5000,
  fry_rel = "BH",
  fry_prod = 0.7,
  fry_capacity = Inf,
  fry_sdev = fry_sdev
)

# Set number of juveniles at beginning of simulation
# Start with 1,000 natural-origin juveniles in each age class
# No hatchery-origin juveniles, i.e., starting a hatchery program
Historical <- new(
  "Historical",
  InitNjuv_NOS = array(1000, c(nsim, Bio@maxage, 1)),
  InitNjuv_HOS = array(0, c(nsim, Bio@maxage, 1))
)

SOM <- new(
  "SOM",
  nsim = nsim,
  proyears = proyears,
  Bio = Bio,
  Habitat,
  Hatchery,
  Harvest,
  Historical
)
```


# Wrapper function

Now let's put the above code chunk into a function. We'll call it `wrapper()`.

This function has three arguments:

- `MSF_vector` a vector that turns on or off mark-selective fishing
- `hatchery_vector` that contains various hatchery production values
- `egg_vector` conrains various egg survival values
- `i` which is an integer that indexes the three previous vectors. When `wrapper()` is called, you supply a value for `i` and the function will grab the `i`-th entry of each vector for subsequent model run.

The rest of the function should be the same as the previous chunk.

It will re-create the same operating model each time the function is called. Each model is identical except for the three management levers.

The wrapper function will then also run the projection:

```{r}
wrapper <- function(i, MSF_vector, hatchery_vector, egg_vector) {

  library(salmonMSE)
  
  nsim <- 20
  proyears <- 30
  maxage <- 3
  
  # Identify management levers
  MSF <- MSF_vector[i]
  hatchery_production <- hatchery_vector[i]
  egg_survival <- egg_vector[i]
  
  # Identify stochastic parameters
  set.seed(234)
  surv1 <- runif(nsim, 0.005, 0.015) # Age 1 marine survival
  surv2 <- 0.8 # Age 2 marine survival is fixed
  
  # Egg-fry survival
  sfry_mean <- 0.7
  sdev_sd <- 0.1
  sdev <- EnvStats::rnormTrunc(nsim * proyears, 1, sdev_sd, min = 0, max = 1/sfry_mean) |> 
    matrix(nsim, proyears)
  
  Mjuv_NOS <- array(, c(nsim, maxage-1, proyears, 1))
  Mjuv_NOS[, 1, , ] <- -log(surv1)
  Mjuv_NOS[, 2, , ] <- -log(surv2)
  Bio <- new(
    Class = "Bio",
    maxage = maxage,
    p_mature = c(0, 0.3, 1),
    #SRrel = "Ricker",
    #kappa = c(3, 4, 6),
    #Smax = 1000,
    #phi = 1,
    #tau = 1,
    Mjuv_NOS = Mjuv_NOS,
    fec = c(0, 2000, 5000),
    p_female = 0.5,
    s_enroute = 0.95
  )
  
  Harvest <- new(
    "Harvest",
    u_preterminal = 0.4,
    u_terminal = 0.2,
    vulPT = c(0, 0.4, 1),
    vulT = c(0, 0.4, 1),
    MSF_PT = FALSE,
    MSF_T = MSF,
    release_mort = c(0.05, 0.05)
  )
  
  Hatchery <- new(
    "Hatchery",
    n_r = 1,
    n_subyearling = hatchery_production,
    s_prespawn = 0.95,
    s_egg_subyearling = 0.7,
    Mjuv_HOS = Bio@Mjuv_NOS,
    p_mature_HOS = Bio@p_mature,
    gamma = 0.8,
    m = 1,
    pmax_esc = 1,
    pmax_NOB = 0.8,
    ptarget_NOB = 0.75,
    phatchery = NA,
    fec_brood = Bio@fec,
    fitness_type = c("none", "none")
  )
  
  Habitat <- new(
    "Habitat",
    use_habitat = TRUE,
    egg_rel = "BH",
    egg_prod = egg_survival,
    egg_capacity = 1000 * 5000,
    fry_rel = "BH",
    fry_prod = sfry_mean,
    fry_capacity = Inf,
    fry_sdev = sdev
  )
  
  # Set number of juveniles at beginning of simulation
  # Start with 1,000 natural-origin juveniles in each age class
  # No hatchery-origin juveniles, i.e., starting a hatchery program
  Historical <- new(
    "Historical",
    InitNjuv_NOS = array(1000, c(nsim, Bio@maxage, 1)),
    InitNjuv_HOS = array(0, c(nsim, Bio@maxage, 1))
  )
  
  SOM <- new(
    "SOM",
    nsim = nsim,
    proyears = proyears,
    Bio = Bio,
    Habitat,
    Hatchery,
    Harvest,
    Historical
  )
  
  SMSE <- salmonMSE(SOM)

  return(SMSE)
}
```

# Creating simulation design

We need to create our simulation design, i.e., which combinations of management levers we will use.

The `expand.grid()` function creates every combination that we specify in the arguments that we supply.
Each row will identify a combination of management levers for our simulation:

```{r}
simulation_design <- expand.grid(
  MSF_vector = c(TRUE, FALSE),
  hatchery_vector = c(1, 2, 3) * 100000,
  egg_vector = c(0.1, 0.15, 0.2)
)
```

# Running wrapper function

We have our wrapper function and simulation design, we can now run the simulation.

There's various ways to run the projections depending on our computation capabilities, e.g., old laptop, new desktop, etc.

For example, we can run a loop over each row of our `simulation_design` data frame or use `lapply()` to facilitate repeated simulations. 

```{r, eval=FALSE}
# Not intended to run, run in parallel instead?
SMSE_list <- list()
for (i in 1:nrow(simulation_design)) {
  SMSE_list[[i]] <- wrapper(
    i, 
    MSF_vector = simulation_design$MSF_vector,
    hatchery_vector = simulation_design$hatchery_vector,
    egg_vector = simulation_design$egg_vector
  )
}
```

```{r, eval=FALSE}
# Not intended to run, run in parallel instead?
SMSE_list <- lapply(
  1:nrow(simulation_design), 
  wrapper, 
  MSF_vector = simulation_design$MSF_vector,
  hatchery_vector = simulation_design$hatchery_vector,
  egg_vector = simulation_design$egg_vector
)
```

But loops are computationally slower compared to parallel processing. We use the `snowfall` package to run simulations over 10 CPUs, then save the results:

```{r run-sim-list, eval=FALSE}
library(snowfall)
cpus <- parallel::detectCores() - 2
sfInit(parallel = TRUE, cpus = cpus)

SMSE_list <- sfLapply(
  1:nrow(simulation_design), 
  wrapper,
  MSF_vector = simulation_design$MSF_vector,
  hatchery_vector = simulation_design$hatchery_vector,
  egg_vector = simulation_design$egg_vector
)

sfStop()

if (!dir.exists(here("output"))) dir.create(here("output"))
saveRDS(SMSE_list, here(file.path("output", "C2-list.rds")))
```

# Results

Now we have a list of `SMSE` objects in `SMSE_list`. 

## Compare time series

Whereas we saw previously that `report()` generates a Markdown report from a single model run, we alternatively now want to compare several model runs simultaneously in a figure.

We have the `compare()` function which generates another markdown report that plots state variables from separate model runs. 

We have 36 scenarios, so we may not be able to neatly all into individual figures. 
So let's compare just the first 3 model runs:

```{r, echo=FALSE}
SMSE_list <- readRDS(here(file.path("output", "C2-list.rds")))
```

```{r}
# Create names for each model run by pasting the SAR and ER values
SMSE_names <- paste0("Hatchery ", simulation_design$hatchery_vector, ", egg surv. = ", simulation_design$egg_vector, 
                     ifelse(simulation_design$MSF_vector, ", No MSF", ", MSF"))
```

```{r, eval=TRUE}
compare(
  SMSE_list[1:3],
  SMSE_names[1:3],
  filename = "C2_compare",
  dir = here("output")
)
```


## Plotting functions

Let's look at PNI, a measure between 0-1 of hatchery influence on the population where higher values indicate lower hatchery influence.

By default, `compare_statevar_ts()` generates a time series figure with base graphics. it generates line plots for each stochastic simulation in the projection (which is messy), so let's directly proceed to quantile figures:

```{r}
compare_statevar_ts(SMSE_list[1:3], var = "PNI", quant = TRUE, names = SMSE_names[1:3])
```

We can also look to distributions within scenarios with `compare_statevar_hist()`, where there are choices to plot using the `density()` or `hist()` functions:

```{r}
compare_statevar_hist(SMSE_list[1:3], var = "PNI", names = SMSE_names[1:3], type = "density")
```

```{r}
compare_statevar_hist(SMSE_list[1:3], var = "PNI", names = SMSE_names[1:3], type = "hist")
```

## Performance measures and decision tables

Risk evaluation typically entails calculating a summary metric that describes the frequency that an objective was met in the simulation. 

For example, we can calculate the probability that PNI exceeded 0.8 (system has high natural influence).
We have a function `PNI80()` that does the calculation for us:

Using the vectorization capabilities in R, we use `sapply()` to call `PNI80()` for each of the objects in `SMSE_list`. We add an extra argument `Yrs = c(30, 30)` that specify the year range (we do the calculation at the end of the projection in year 30:

```{r}
sim_results <- simulation_design
sim_results$PNI_objective <- sapply(SMSE_list, PNI80, Yrs = c(30, 30))
```

We want to explore the decision space: what combinations of levers help us get high PNI and which levers are more influential than others?

`plot_decision_table()` helps us plot a grid of values using ggplot.
The colour scheme maps 0-1 with red to green (0.5 is white).

We see a clear delineation where we don't meet our high PNI objective when egg survival is too low. Marginally, high hatchery production hurts us as well. 
On the other hand, mark-selective fishing helps with PNI management.

```{r}
sim_results$MSF_label <- ifelse(sim_results$MSF_vector, "MSF", "No MSF")
plot_decision_table(
  x = sim_results$hatchery_vector,
  y = sim_results$egg_vector,
  z = sim_results$PNI_objective,
  xlab = "Target releases",
  ylab = "Egg incubation survival",
  title = "Probability PNI > 0.8",
  scenario = sim_results$MSF_label
)
```

Alternatively, `plot_decision_table2()` gives us more flexibility in the colour scheme by binning the summary metric into separate colour bins. Using the above figure, we refine the decision table based on the values we see in our results.

```{r}
plot_decision_table2(
  x = sim_results$hatchery_vector,
  y = sim_results$egg_vector,
  z = sim_results$PNI_objective,
  xlab = "Target releases",
  ylab = "Egg incubation survival",
  title = "Probability PNI > 0.8",
  scenario = sim_results$MSF_label,
  bin = c(0, 0.5, 0.7),
  bin_labels = c("0.00 - 0.49", "0.50 - 0.69", "0.70 - 1"),
  bin_col = c("pink", "green", "green4"),
  add_values = TRUE
)
```

We can envision running a larger grid to map out the full decision space, or even adding an additional management lever to our analysis. `plot_decision_table()` can readily handle up to four levers: we have the x and y axes of a table (two levers), as well as panels of tables in row and columns (two more levers).

If our decision making context is solely on PNI management, then these figure can be the primary result from a simulation analysis. (Of course, other factors can come into play as to whether we can move the needle to increase egg survival, implement mark-selective fishing or not. This is an artifical example, but there are usually motivating reasons that help guide the direction of a simulation analysis. We also will likely know what is possible and unlikely on the management side.)


## Tradeoffs

Frequently, objectives are conflicting. For example, higher hatchery production should increase natural spawner abundance at the cost of PNI. Tradeoff figures allow us to compare two summary metrics simultaneously.

Let's compare median PNI and median natural spawners (sum of natural-origin and hatchery-origin spawners, excluding brood) in the long-term (end of the projection).

First, let's calculate our second metric, average catch, with a custom function:

```{r}
PNI_fn <- function(SMSE) median(SMSE@PNI[, 1, 30])
Spawner_fn <- function(SMSE) {
  Spawner <- SMSE@NOS + SMSE@HOS
  Spawner_total <- rowSums(Spawner[, 1, , 30])
  median(Spawner_total)
}
sim_results$PNI <- sapply(SMSE_list, PNI_fn)
sim_results$Spawners <- sapply(SMSE_list, Spawner_fn)
```

Next, we can use `plot_tradeoff()` to plot the two metrics (PNI and spawners) against each other.

We see a clear frontier lines where increasing hatchery production (red to green to blue) decreases PNI and increases spawners.
The frontier line also moves upward and to the right (better outcomes for both PNI and spawners) with improved egg survival in the natural environment and with mark-selective fishing.

Between the decision tables and tradeoff figure, which gives the clearer take-home message?

```{r}
library(ggplot2)
plot_tradeoff(
  pm1 = sim_results$Spawners,
  pm2 = sim_results$PNI,
  x1 = factor(sim_results$hatchery_vector),
  x2 = sim_results$egg_vector,
  xlab = "Median Spawners",
  ylab = "Median PNI",
  x1lab = "Hatchery production",
  x2lab = "Egg incubation survival",
  scenario = sim_results$MSF_label
) +
  scale_shape_manual(values = c(1, 4, 16))
```




